---
title: "ICA 5"
author: "Jonathan Thompson"
date: "2024-03-18"
output: pdf_document
---

## Customer Churn

In this assignment, you will estimate a survival analysis model on customer churn. In this dataset, you will find various characteristics about a set of phone company customers

customerID: Customer ID number
gender: Provides the stated gender of the customer
SeniorCitizen: States if the person is a senior citizen
Partner: Does the person have a partner
Dependents: Does the person have dependents
tenure: States how long the person has been with the Bank
PhoneService: Do they have phone service
MultipleLines: Do they have multiple lines
InternetService: What type of Internet Service do they have (DSL, fiber optic, none)
OnlineSecurity: Do they have online security?
OnlineBackup: Do they have online backup?
DeviceProtection: Do they have device protection?
TechSupport: Did they use tech support?
StreamingTV: Do they use the internet to stream TV?
StreamingMovies: Do they use the internet to stream movies?
Contract: What type of contract do they have (month-to-month, one-year, or two year)
PaperlessBilling: Do they use paperless biling
PaymentMethod: How do they pay for services
MonthlyCharges: What is their monthly charge?
TotalCharges: What is their total charge for the quarter?
Churn: Did they leave the company


1) Find the simple average of `tenure`. Explain why this simple average can be biased.

```{r warning=FALSE, message=FALSE, eval=TRUE}

library(readr)

Churn <- read.csv ("C:/Users/catho/OneDrive/Documents/AAT/WA_Fn-UseC_-Telco-Customer-Churn.csv")


```

Answer: The average number of months a customer remains with the bank is  `r  round (mean(Churn$tenure), digits=2)` months. In survival analysis, it often contains censored observations that have not occurred by the end of a certain study or experiment. If a simple average is used without accounting for censoring, it may produce biased results. 

2) Find the simple average of `tenure` by the following groups. Gender, SeniorCitizen, and Partner.

```{r eval=TRUE}

Churn$SeniorCitizen <- factor(Churn$SeniorCitizen, labels= c("Non-Senior", "Senior"))

tab1 <- aggregate(Churn$tenure, by=list(Churn$gender, Churn$SeniorCitizen, Churn$Partner), FUN=mean)

names(tab1) <-c ("Gender", "Senior Status", "Partnered", "Mean of Tenure")

knitr::kable(tab1)

library(modelsummary)

datasummary(tenure ~ Mean*(Gender=gender)+Mean*(`Senior Citizen`= SeniorCitizen)+Mean*(Partner), data=Churn)


```


3) Find the simple average of `tenure` and `MonthlyCharge` by `Contract` type.
```{r eval=TRUE}

datasummary((`Tenure`=tenure)+(`Monthly Charge`= MonthlyCharges)~Mean*Contract, data=Churn)


```


4) Estimate a Kaplan Meier survival model. Use gender and senior citizen as explanatory variables. Do these variables produce statistically different survival rates?

```{r eval=TRUE}

Churn$delta <-ifelse(Churn$Churn=="Yes", 1, 0)

survminer::ggsurvplot(
    fit = survival::survfit(survival::Surv(tenure, delta) ~ gender+SeniorCitizen, data = Churn), 
    xlab = "Months",
    ylab = "Overall survival probability",
    legend.title = "Customer Type",
    legend.labs = c("Female & Non-Senior", "Female & Senior", "Male & Non-Senior","Male & Senior"),
    break.x.by = 100, 
    censor = FALSE,
    conf.int = TRUE)

```
Answer: Yes, but they do not specifically differ in regards to gender. They differ in regards to if they are a Senior or Non-Senior. Both Male and Female Non-Seniors are at about the same survival probability rate at around .69 while Male & Female Seniors at at a much lower rate of about .50. As the line graph ends, both drop lower and lower. However, the Seniors of both genders drop at a much lower rate. 


5) Estimate a Kaplan Meier survival model that uses contract type as the explanatory variable. Do we see a difference between contract type?

```{r eval=TRUE}

Churn$delta <-ifelse(Churn$Churn=="Yes", 1, 0)

survminer::ggsurvplot(
    fit = survival::survfit(survival::Surv(tenure, delta) ~ Contract, data = Churn), 
    xlab = "Months",
    ylab = "Overall survival probability",
    legend.title = "Customer Type",
    legend.labs = c("Month to Month", "One Year", "Two Year"),
    break.x.by = 100, 
    censor = FALSE,
    conf.int = TRUE)


```
Answer: Yes, especially with the Month-To-Month contract type. Within this contract type, it instantly drops within the first month, gradually gets lower & ends at around a probability of .15. The One-Year contract type drops mildly lower as each month accrues but drops towards the very end month(s) at around .60. The Two-Year contract type has no difference in decline until they very end month(s) in the upper .90's.


6) Estimate a Cox proportional hazard model of `tenure`. Use the following variables as explanatory variables: gender, Senior Citizen, contract type, partner, dependents, type of internet access, do they have phone service, and do they use paperless billing.

```{r eval=TRUE}

library(broom)
library(survival)

model_cox <- survival::coxph(
  survival::Surv(tenure, delta) ~ gender + SeniorCitizen + Contract + Partner + Dependents + InternetService + PhoneService + PaperlessBilling,
  data = Churn)

tidy_broom <- broom::tidy(model_cox)

coef_names <- c("Male", "Senior Citizen", "One-Year Contract", "Two-Year Contract", "Yes To Partner", "Yes To Dependents", "Optic Internet Service", "No To Internet Service", "Yes To Phone Service", "Yes To Paperless Billing")

tidy_broom$term <- coef_names

colnames(tidy_broom)[2:5] <- c("Estimate", "Standard Error", "Statistic", "P-Value")

print(tidy_broom)



```

