---
title: "My First R Markdown Document"
author: "Nick Holt"
date: "December 5, 2020"
output: html_document
---


## Two hashes denote a new section
Any text that you type in a normal way will be displayed as plain text. This is where you might write the body of your document.

> To offset text in a block quote simply use the greater than symbol before the line of text.
  
> Including a double space will offset even more.

*Italics are added by including the asterisk symbol before and after your text.*

Bullets can be added as well.

* Use an asterisk, and a space to add bullets
* **emphasis** can be added by including two asterisks:
  
  * and sub bullets can be added with a double space and asterisk
  
  * like so

You can add subscript by using math notation (place $ before and after the text)
Here is a subscript example: 

* $H_0:$ R Markdown is amazing
* $H_1:$ R Markdown is scary and difficult


```{r}

```


ctrl + alt + i will add a chunk of code automatically

```{r}
plot(iris)
```

R markdown can embed your output seamlessly with your code and your writeup.

Let's make the plot smaller by adding fig.width and fig.height to our code chunk

```{r fig.width = 5, fig.height = 5.5}
plot(iris)
```

Here is a good [resource](www.youtube.com/watch?v=LIIC8cViC54)


There is a lot more we can learn. Anything you need can be found with a quick google search.